---
output_dir: "."
title: "RNotebookExample"
output: html_document
---

```{r load-data and tidy-dataset, message=FALSE}
#load library
library(tidyverse)
library(kableExtra) #used fro the kbl()
library(rmarkdown)
library(ggpubr) #used for ggarrange()

#load the file and save it in an object
pop_data <- read_csv("data/gapminder_data.csv")

# tranform your data to shorten the file
# creata a table with data for european countries in 2007 - ordered by life expectancy

euro_data_tbl <- pop_data %>% 
  filter(continent == "Europe" & year == 2007) %>% 
  select(-continent, -year) %>% 
  arrange(desc(lifeExp)) %>% 
  rename(Country = country, "Population sze" = pop, "Life expectancy" = lifeExp, "GDP" = gdpPercap)

```


o show results in a table we use the kbl() function from the [library(kableExtra)](https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html)

The results in euro_data_tbl are dsplayed in the table below:

```{r kbl}
euro_data_tbl %>% 
  kbl() %>% 
  kable_styling(bootstrap_options = c("striped", "hover"))



```


If you have a very long table to display, the best way would be to use pagination where the contents of the table are split into multiple tabs. The function paged_table in the rmarkdown library can be used to do this. It is also very simple to use.

Create a new R code chunk called paged-table and use the paged_table() function to display the euro_data_tbl table as follows:

```{r paged-table}
paged_table(euro_data_tbl)

```


# Adding images



Next, the life expectancy in Europe is observed across different years.  The aim of this is to check if there was a change in life expectancy over the years.  The data used for the life expectancy over the years analysis is the gapminder data and the following data manipulation was performed:

```{r fig-data}
#keep on European data and change year to factor
euro_data_fig <- pop_data %>% 
  filter(continent == "Europe") %>% 
  mutate(year=as_factor(year))


```

The life expectancy for Europe over the years is plotted as following:


```{r fig, warning=FALSE, fig.cap="Life Expectancy in Europe over the years"}
#Euro plot
euro_plot <- euro_data_fig %>%
                ggplot(mapping=aes(x=year, y=lifeExp)) +
                geom_violin() +
                stat_summary(fun.y = median, geom = "point")

#draw euro plot
euro_plot

```


# Make images with multiple panels


```{r fig with panels}
#Uk data fig

uk_data_fig <- pop_data %>% 
  filter(country == "United Kingdom") %>% 
  mutate(year=as_factor(year))

# fig-composition

uk_plot <- uk_data_fig %>% 
  ggplot(aes(x=year, y=lifeExp))+
  geom_point()

ggarrange(euro_plot, uk_plot, ncol = 2, nrow = 1, labels = "AUTO")

```




